---
author: "University of Melbourne Centre for Cancer Research"
date: "`r format(Sys.time(), '%a %Y-%b-%d %H:%M')`"
output:
  html_document:
    theme: simplex
    toc: true
    toc_float: false
    toc_depth: 5
    code_download: true
    code_folding: hide
  rmdformats::material:
    highlight: kate
params:
  woof_final: "/path/to/woof/final"
  run1_nm: "run1"
  run2_nm: "run2"
title: "`r glue::glue('Comparison of woof runs in {htmltools::br()}\'{normalizePath(params$woof_final)}\':{htmltools::br()}{params$run1_nm}{htmltools::br()}  vs.{htmltools::br()}{params$run2_nm}')`"
description: "Comparison of bcbio and umccrise pipeline runs"
---

<style type="text/css">
.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r load_knitr, include=F}
require(knitr)
knitr::opts_chunk$set(
  echo = FALSE
  )
```

```{r render_report_interactively, eval=F, echo=F}
params_tmp <- list(
  local = list(
    woof_final = "/Users/pdiakumis/Desktop/projects/umccr/woof_compare/test/woof/final",
    run1_nm = "run1",
    run2_nm = "run2"
  ),
  gadi = list(
    woof_final = "",
    run1_nm = "run1",
    run2_nm = "run2"
  )
)
params <- params_tmp[["local"]]

rmd <- here::here("inst/rmd/woof_compare.Rmd")
render_me <- function() {
  rmarkdown::render(
    rmd,
    params = params)
}
render_me()
```

```{r load_pkgs, message=F, warning=F}
suppressPackageStartupMessages(library(bedr))
suppressPackageStartupMessages(library(DiagrammeR))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(glue))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(rock))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(woofr))
```

Results
-------

### SNVs/Indels

```{r funcs_snvs}
woof_final <- params$woof_final
stopifnot(dir.exists(woof_final))

samples <- list.dirs(woof_final, recursive = FALSE, full.names = FALSE)
# get vcf_types from unique woof/final/<sample>/vcf_eval/<vcf_type>/
snv_vcf_types <-
  file.path(woof_final, samples, "vcf_eval") %>%
  list.dirs(recursive = FALSE, full.names = FALSE) %>%
  unique()
# allowed_snv_vcf_types <- c("ensemble-batch", "ensemble-germ", "gatk-germ", "mutect-batch",
#                            "strelka-batch", "strelka-germ", "vardict-batch", "vardict-germ",
#                            "pcgr", "cpsr")
# stopifnot(snv_vcf_types %in% allowed_snv_vcf_types)

read_count_files <- function(woof_final, sample, vcf_type, pass_or_all) {

  c1 <- file.path(woof_final, sample, "vcf_counts", "f1", vcf_type, pass_or_all, "count_vars.txt") %>% woofr::read_count_file()
  c2 <- file.path(woof_final, sample, "vcf_counts", "f2", vcf_type, pass_or_all, "count_vars.txt") %>% woofr::read_count_file()

  # keeping sample/vcf_type/subset in row
  dplyr::tibble(sample = sample, vcf_type = vcf_type, subset = pass_or_all, run1_count = c1, run2_count = c2)
}

get_stats <- function(woof_final, sample, vcf_type, pass_or_all) {
  dplyr::bind_cols(
    read_count_files(woof_final, sample, vcf_type, pass_or_all),
    woofr::read_eval_file(file.path(woof_final, sample, "vcf_eval", vcf_type, pass_or_all, "eval_stats.tsv")))
}

get_stats_sample <- function(sample, vcf_types) {

  get_stats_sample_subset <- function(sample, pass_or_all) {
    vcf_types %>%
      purrr::map(~ get_stats(woof_final, sample, .x, pass_or_all)) %>%
      dplyr::bind_rows()
  }

  dplyr::bind_rows(
    get_stats_sample_subset(sample, "PASS"),
    get_stats_sample_subset(sample, "ALL")
  )
}
```

```{r analysis_snvs}
d <- purrr::map(samples, ~ get_stats_sample(.x, snv_vcf_types)) %>%
  dplyr::bind_rows() %>%
  dplyr::mutate(id = dplyr::row_number())
```

```{r tab_descr_snvs}
x <- dplyr::tribble(
  ~Column,    ~Description,
  "run1",     "woof run1",
  "run2",     "woof run2 (truthset)",
  "subset",   "All/PASSed variants",
  "SNP",      "SNP results",
  "IND",      "Indel results",
  "FP",       "False Positive",
  "FN",       "False Negative",
  "TP",       "True Positive",
  "Truth",    "TP + FN",
  "Recall",   "TP / Truth",
  "Precision", "TP / TP + FP"
)

cap <- x %>%
  mutate(y = paste(Column, Description, sep = ": ")) %>%
  pull(y)
cap <- htmltools::tags$caption(htmltools::div(paste(cap[1:5], collapse = "; ")),
                               htmltools::div(paste(cap[6:length(cap)], collapse = "; ")))
```

#### Comparison Table

```{r eval_tab_snvs}
bgsize <- '90% 90%'
big_mark_cols <- c("run1_count", "run2_count",
                   "SNP_Truth", "SNP_TP", "SNP_FP", "SNP_FN",
                   "IND_Truth", "IND_TP", "IND_FP", "IND_FN")
bar_colour <- "lightblue"
bar_range <- c(0.97, 1)

d %>%
  dplyr::select(-matches("f[123]$")) %>%
  dplyr::select(id, sample, vcf_type, subset, run1_count, run2_count,
                contains("Recall"), contains("Precision"),
                everything()) %>%
  dplyr::mutate_if(is.numeric, round, 3) %>%
  dplyr::mutate_if(is.character, as.factor) %>%
  DT::datatable(
    rownames = FALSE, caption = cap, class = "cell-border stripe hover compact",
    filter = list(position = "top", clear = FALSE, plain = FALSE),
    extensions = c("Scroller", "KeyTable"),
    options = list(scroller = TRUE, scrollY = 600, autoWidth = TRUE, keys = TRUE)
  ) %>%
  formatStyle(
    'SNP_Recall',
    background = styleColorBar(bar_range, color = bar_colour),
    backgroundSize = bgsize, backgroundRepeat = 'no-repeat', backgroundPosition = 'center'
  ) %>%
  formatStyle(
    'IND_Recall',
    background = styleColorBar(bar_range, color = bar_colour),
    backgroundSize = bgsize, backgroundRepeat = 'no-repeat', backgroundPosition = 'center'
  ) %>%
  formatStyle(
    'SNP_Precision',
    background = styleColorBar(bar_range, color = bar_colour),
    backgroundSize = bgsize, backgroundRepeat = 'no-repeat', backgroundPosition = 'center'
  ) %>%
  formatStyle(
    'IND_Precision',
    background = styleColorBar(bar_range, color = bar_colour),
    backgroundSize = bgsize, backgroundRepeat = 'no-repeat', backgroundPosition = 'center'
  ) %>%
  formatCurrency(big_mark_cols, currency = "", interval = 3, mark = ",", digits = 0)
```

### Structural Variants

```{r funcs-svs}
read_manta_both <- function(sample, f1, f2) {

  read_manta_file <- function(sample, f1_or_f2) {
    stopifnot(sample %in% samples, f1_or_f2 %in% c(f1, f2))
    manta_fnames <- list.files(file.path(raijin_share, sample, f1_or_f2),
                               pattern = "prioritize.*vcf.gz$", recursive = TRUE, full.names = TRUE)
    manta_som_fname <- manta_fnames[!grepl("-germline-sv-prioritize", basename(manta_fnames))]
    stopifnot(length(manta_som_fname) == 1)

    sv <- manta_som_fname %>%
      normalizePath() %>%
      rock::prep_manta_vcf()

    list(sv = sv$sv, fname = manta_som_fname)
  }

  sv1 <- read_manta_file(sample, f1)
  sv2 <- read_manta_file(sample, f2)

  list(sv1 = list(sv = sv1$sv, fname = sv1$fname),
       sv2 = list(sv = sv2$sv, fname = sv2$fname))
}

manta_isec <- function(sample, f1, f2, bnd_switch = TRUE) {

  switch_bnd <- function(d) {
    stopifnot(all(colnames(d) %in% c("chrom1", "pos1", "chrom2", "pos2", "svtype")))
    no_bnd <- dplyr::filter(d, !.data$svtype %in% "BND")
    bnd <-
      dplyr::filter(d, .data$svtype %in% "BND") %>%
      dplyr::select(chrom1 = .data$chrom2, pos1 = .data$pos2,
                    chrom2 = .data$chrom1, pos2 = .data$pos1, .data$svtype)

    dplyr::bind_rows(no_bnd, bnd)
  }

  mb <- read_manta_both(sample, f1, f2)
  manta_fname1 <- mb$sv1$fname
  manta_fname2 <- mb$sv2$fname
  tot_vars <- list(run1 = nrow(mb$sv1$sv), run2 = nrow(mb$sv2$sv))
  col_nms <- colnames(mb$sv1$sv)
  fp <- dplyr::anti_join(mb$sv1$sv, mb$sv2$sv, by = col_nms)
  fn <- dplyr::anti_join(mb$sv2$sv, mb$sv1$sv, by = col_nms)

  if (bnd_switch) {
    # some real matching cases simply have switched BND mates
    fp_new <- dplyr::anti_join(switch_bnd(fp), fn, by = col_nms) %>% switch_bnd()
    fn_new <- dplyr::anti_join(switch_bnd(fn), fp, by = col_nms) %>% switch_bnd()
    fp <- fp_new
    fn <- fn_new
  }

  return(list(sample = sample,
              file1 = manta_fname1,
              file2 = manta_fname2,
              tot_vars = tot_vars,
              fp = fp,
              fn = fn))
}

manta_isec_stats <- function(mi) {
  samplename <- mi$sample
  fn <- nrow(mi$fn)
  fp <- nrow(mi$fp)
  tp <- mi$tot_vars$run2 - fn
  truth <- tp + fn
  recall <- tp / truth
  precision <- tp / (tp + fp)

  dplyr::tibble(
    sample = samplename, vcf_type = "manta-svpri",
    run1_count = mi$tot_vars$run1, run2_count = mi$tot_vars$run2,
    Recall = round(recall, 3), Precision = round(precision, 3), Truth = truth,
    TP = tp, FP = fp, FN = fn
  )
}

get_circos <- function(mi) {

  prep_svs_circos <- function(mi) {
    sv <- dplyr::bind_rows(list(fp = mi$fp, fn = mi$fn), .id = "fp_or_fn")
    links_coloured <- sv %>%
      dplyr::mutate(chrom1 = paste0("hs", .data$chrom1),
                    chrom2 = paste0("hs", .data$chrom2)) %>%
      dplyr::mutate(col = dplyr::case_when(
        fp_or_fn == "fp" ~ '(0,255,0)',
        fp_or_fn == "fn" ~ '(255,0,0)',
        TRUE ~ '0,0,0')) %>%
      dplyr::mutate(pos1b = .data$pos1,
                    pos2b = .data$pos2,
                    col = paste0('color=', col)) %>%
      dplyr::select(.data$chrom1, .data$pos1, .data$pos1b, .data$chrom2, .data$pos2, .data$pos2b, .data$col)
    return(links_coloured)
  }

  write_circos_configs <- function(mi) {
    sample <- mi$sample
    outdir <- file.path(raijin_share, sample, "woof/final/circos/manta_isec")
    dir.create(outdir, recursive = TRUE)

    links <- prep_svs_circos(mi)
    message(glue::glue("Writing circos links file to '{outdir}'."))
    readr::write_tsv(links, file.path(outdir, "SAMPLE.link.circos"), col_names = FALSE)

    message(glue::glue("Copying circos templates to '{outdir}'"))
    file.copy(from = system.file("templates/circos", glue::glue("circos_sv.conf"), package = "pebbles"),
              to = file.path(outdir, "circos.conf"), overwrite = TRUE)
    file.copy(system.file("templates/circos", "gaps.txt", package = "pebbles"), outdir, overwrite = TRUE)
    file.copy(system.file("templates/circos", "ideogram.conf", package = "pebbles"), outdir, overwrite = TRUE)

    return(outdir)
  }

  run_circos <- function(outdir, sample) {
    export_path <- "export PATH=/g/data3/gx8/extras/umccrise/miniconda/envs/umccrise_purple/bin:/g/data3/gx8/extras/umccrise/miniconda/bin:/opt/bin:/bin:/usr/bin:/opt/pbs/default/bin"
    cmd <- glue::glue("{export_path} && echo $(date) running circos on {sample} && circos -nosvg -conf {outdir}/circos.conf -outputdir {outdir} -outputfile circos_{sample}.png")
    system(cmd, ignore.stdout = TRUE)
    return(file.path(outdir, glue::glue("circos_{sample}.png")))
  }

  outdir <- write_circos_configs(mi)
  circos_png <- run_circos(outdir, mi$sample)
  circos_png
}


sv_results_per_sample <- function(sample) {
  mi <- manta_isec(sample, f1, f2)
  mi_stats <- manta_isec_stats(mi)
  circos <- get_circos(mi)

  list(mi = mi, stats = mi_stats, circos = circos)
}
```




```{r analysis-svs, message=F, warning=F}
sv_results <-
  purrr::map(samples, sv_results_per_sample) %>%
  purrr::set_names(samples)
```

```{r tab-descr-svs}
x <- dplyr::tribble(
  ~Column,    ~Description,
  "run1",     f1,
  "run2",     paste(f2, "(truthset)"),
  "FP",       "False Positive",
  "FN",       "False Negative",
  "TP",       "True Positive",
  "Truth",    "TP + FN",
  "Recall",   "TP / Truth",
  "Precision", "TP / TP + FP"
)

cap <- x %>%
  mutate(d = paste(Column, Description, sep = ": ")) %>%
  pull(d)
cap <- htmltools::tags$caption(htmltools::div(paste(cap[1:4], collapse = "; ")),
                               htmltools::div(paste(cap[5:length(cap)], collapse = "; ")))
```


#### Comparison Table

```{r eval-tab-svs}
bgsize <- '90% 90%'
big_mark_cols <- c("run1_count", "run2_count",
                   "Truth", "TP", "FP", "FN")

sv_stats <- purrr::map_dfr(sv_results, "stats")
sv_stats %>%
  dplyr::mutate_if(is.character, as.factor) %>%
  DT::datatable(
    rownames = FALSE, caption = cap,
    filter = list(position = "top", clear = FALSE, plain = TRUE),
    extensions = c("Scroller"),
    options = list(scroller = TRUE, scrollX = TRUE, scrollY = 220)) %>%
  formatStyle(
    'Recall',
    background = styleColorBar(sv_stats$Recall, 'lightgreen'),
    backgroundSize = bgsize,
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  ) %>%
  formatStyle(
    'Precision',
    background = styleColorBar(sv_stats$Precision, 'lightgreen'),
    backgroundSize = bgsize,
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  ) %>%
  formatCurrency(big_mark_cols, currency = "", interval = 3, mark = ",", digits = 0)
```

#### FP/FN Table
Showing FP and FN SVs from above:

```{r sv-fp-fn-table}
fp <- purrr::map(sv_results, function(x) x[["mi"]][["fp"]]) %>%
  dplyr::bind_rows(.id = "sample")
fn <- purrr::map(sv_results, function(x) x[["mi"]][["fn"]]) %>%
  dplyr::bind_rows(.id = "sample")
fp_fn <- dplyr::bind_rows(list(fp = fp, fn = fn), .id = "FP_or_FN")

fp_fn %>%
  dplyr::mutate_if(is.character, as.factor) %>%
  DT::datatable(
    rownames = FALSE,
    filter = list(position = "top", clear = FALSE, plain = TRUE),
    extensions = c("Scroller"),
    options = list(scroller = TRUE, scrollX = TRUE, scrollY = 400)) %>%
  formatCurrency(c("pos1", "pos2"), currency = "", interval = 3, mark = ",", digits = 0)
```


#### Circos Plots {.tabset .tabset-fade}

* <span style="color:green">Green</span> = False Positives
* <span style="color:red">Red</span> = False Negatives

```{r prep-sv-circos}
circos_plots <- purrr::map_chr(sv_results, "circos")
```

##### `r i <- 0; samples[i <- i + 1]`

```{r out.width="50%"}
knitr::include_graphics(circos_plots[i])
```

##### `r  samples[i <- i + 1]`

```{r out.width="50%"}
knitr::include_graphics(circos_plots[i])
```

##### `r  samples[i <- i + 1]`

```{r out.width="50%"}
knitr::include_graphics(circos_plots[i])
```

##### `r  samples[i <- i + 1]`

```{r out.width="50%"}
knitr::include_graphics(circos_plots[i])
```

##### `r  samples[i <- i + 1]`

```{r out.width="50%"}
knitr::include_graphics(circos_plots[i])
```



Methods
-------
Here we're comparing the output from two bcbio runs:

- __Run1__: `r params$run1_nm`
- __Run2__ (_"truthset"_): `r params$run2_nm`

* The comparison of VCFs has been done using the `compare` module of [woof](https://github.com/umccr/woof).
* The `Rmd` template used to generate this report is available in [woofr](https://github.com/umccr/woofr).

### SNVs/INDELs

- split into PASS and ALL variants
- run `bcftools isec` to generate VCFs containing variants that are:
  - False Positives (FPs): within Run1, but not Truthset
  - False Negatives (FNs): within Truthset, but not Run1
- True Positives (TPs): within _both_ Run1 and Truthset
- run `eval_vcf` function in <https://github.com/umccr/vcf_stuff/>
  to generate a summary table with evaluation statistics

#### Diagram

```{r graph1, fig.height=7}
set.seed(42)

ndf <-
  tribble(
    ~id, ~label, ~type,
    1, "snv1", "file-initial",
    2, "snv2", "file-initial",
    3, "bcftools-view-f", "command",
    4, "vcf1_pass", "file",
    5, "vcf2_pass", "file",
    6, "bcftools-isec", "command",
    7, "FP-FN-TP", "file",
    8, "FP-FN-TP_pass", "file",
    9, "count-vars", "command",
    10, "counts", "file",
    11, "counts_pass", "file",
    12, "eval-vcf", "command") %>%
  mutate(
    shape = case_when(
      type == "file" ~ "rectangle",
      type == "command" ~ "circle",
      TRUE ~ "square"
    ),
    fillcolor = case_when(
      type == "file" ~ "lightblue",
      type == "command" ~ "#f48f42",
      TRUE ~ "#eef442"
    ),
    fontcolor = "black")

edf <-
  create_edge_df(
    from = c(1, 2, 3, 3, 1, 2, 4, 5, 6, 6, 1, 2, 4, 5, 9, 9, 7, 8, 10, 11),
    to =   c(3, 3, 4, 5, 6, 6, 6, 6, 7, 8, 9, 9, 9, 9, 10, 11, 12, 12, 12, 12)
  )


create_graph(nodes_df = ndf, edges_df = edf) %>%
  set_node_attrs(
    node_attr = "fontsize",
    values = "8"
  ) %>%
  render_graph()
```

### Structural Variants

- keep all variants from Run1 and Run2 (i.e. don't filter any variants)
- generate
  - number of SVs in Run1
  - number of SVs in Run2
  - FP variants: within Run1, but not Truthset
  - FN variants: within Truthset, but not Run1
- generate one circos plot per sample, where:
  - FN in red
  - FP in green

### Copy Number Variants

- check if gene number and coordinates are identical between runs
- check which genes have a difference in Min/Max copy number greater than 0.5
